NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-30 paper/application.tex
finsihed the full draft
space saving
length fit
update application 6.1

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \vspace{-4mm}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples). The user can drill down to individual examples by filtering via the histogram or scatterplot (see Fig.~\ref{fig:summaryView}). For highly unstable prediction, we often observe the original prediction is near a decision boundary (e.g., the yellow circle in (e) corresponds to a prediction that is at the boundary between \emph{entailment} and \emph{neutral}), however, some predictions, such as the one illustrated in (d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word \textbf{pile} is replaced by \textbf{heap} in the hypothesis sentence).
}
\label{fig:predictStability}
\vspace{-2mm}
\end{figure}

@@

Besides presenting the summary view, the tool also allows the user to quickly narrow down the selection to a single example by filtering via the histogram and scatterplot (see Fig.~\ref{fig:summaryView}).
%
Through the exploration of many samples with low stabilities (i.e., high sensitivities values in the plots), domain experts notice that many highly unstable outliers are from sentence pairs where the predictions are near the decision boundary (see Fig.~\ref{fig:predictStability}(e), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
However, some predictions, such as the one illustrated in Fig.~\ref{fig:predictStability}(d) can alter the prediction quite drastically with minor perturbation (shown in Fig.~\ref{fig:predictStability}$d_1$, where the word \textbf{heap} is replaced by \textbf{pile} in the hypothesis sentence). In the following section, we will examine what happened inside the model and hypothesize the cause of failure (see Fig.~\ref{fig:att2pred}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% perturb input, attention
\subsection{Scenario 2: Examine the Decision Making Process}
The predicted label alone provides limited information. Often, domain experts want to know how does the model arrive at a conclusion. And if the prediction is incorrect, where in the model does the error occur?
Examining the decision making process is not only instrumental in evaluating the model performance but also essential for hypothesizing improvement strategies for future models.
%
In the NLI model, the three stages (encoder, attention, classifier) work in synergy to produce the prediction.
Therefore, making sense of the prediction involves understanding how different parts of the model affect the final prediction.

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{att2pred}
 \caption{
Editing the original attention (a) to correctly align the word "heap" with "pile" as shown in (b) (these two words are highlighted in orange).
The change of attention leads to the change of prediction from neutral to the class boundary between \emph{neutral} and \emph{entitlement}.
%
}
\label{fig:att2pred}
\end{figure}

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipelineUpdate}
 \vspace{-2mm}
 \caption{
Experiment with all configurations for the label reassignment optimization. As shown in (d), the update of attention stage seems to have significantly less impact on the prediction result compared to the classifier or encoder of the model.
 }
\label{fig:pipelineUpdate}
\end{figure*}

@@

%
%However, to understand how the model should
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{depTreeExample}
 \caption{
Dependency tree provides valuable information that can help fix the prediction error.
In (a), the model mistakenly aligns the word green, which lead to wrong prediction.  
After examining the dependency tree (highlighted by pink squares), we can see the two \textbf{green} are attached to different words.
In (b), by editing the attention and forcing the two \textbf{green}'s alignment to be zero, the prediction label is corrected to \emph{neutral}. 
 }
\label{fig:depTreeExample}
\end{figure*}

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \vspace{-6mm}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type of predictions over the entire development set (10k examples). The user can drill down to individual examples by using the interface described in Fig.~\ref{fig:summaryView}. In (d), we illustrate a highly unstable prediction, where the original pair's prediction is near a decision boundary (i.e., the yellow circle is between \emph{entailment} and \emph{neutral}). In (e), we show another example of unstable predictions, in which the prediction changes drastically with a minor perturbation (see ($\mathbf{e_1}$)). %( For the perturbed pair , the word \textbf{pile} is replaced by \textbf{heap} in the hypothesis sentence).
}
\label{fig:predictStability}
\vspace{-3mm}
\end{figure}

@@

@@

Besides presenting the summary view, the tool also allows the user to quickly narrow down the selection to a single example by filtering via the histogram and scatterplot (see details in Fig.~\ref{fig:summaryView}).
%
Through the exploration of many samples with low stabilities (i.e., high sensitivities values in the plots), domain experts notice that many highly unstable outliers are from sentence pairs where the predictions are near the decision boundary (see Fig.~\ref{fig:predictStability}(d), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
%
However, we can also find sentence pairs, such as the one illustrated in Fig.~\ref{fig:predictStability}(e), in which the prediction is altered drastically with minor perturbations (e.g., replace the word \textbf{pile} with \textbf{heap} in ``pile of snow''). In the following section, we examine what happened inside the model and hypothesize the cause of failure (see Fig.~\ref{fig:att2pred}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{att2pred}
 \vspace{-6mm}
 \caption{
Editing the original attention (a) to correctly align the word "heap" with "pile" as shown in (b) (these two words are highlighted in orange).
The change of attention leads to the change of prediction from neutral to the class boundary between \emph{neutral} and \emph{entitlement}.
%
}
\label{fig:att2pred}
\end{figure}

% perturb input, attention
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipelineUpdate}
  \vspace{-6mm}
 \caption{
Experiment with all configurations for the label reassignment optimization. As shown in (d), the update of attention stage seems to have significantly less impact on the prediction result compared to the classifier or encoder of the model.
 }
  \vspace{-4mm}
\label{fig:pipelineUpdate}
\end{figure*}

\subsection{Scenario 2: Examine the Decision Making Process}
The predicted label alone provides limited information. Often, domain experts want to know how does the model arrive at a conclusion. And if the prediction is incorrect, where in the model does the error occur?
Examining the decision making process is not only instrumental in evaluating the model performance but also essential for hypothesizing improvement strategies for future models.
%
In the NLI model, the three stages (encoder, attention, classifier) work in synergy to produce the prediction.
Therefore, making sense of the prediction involves understanding how different parts of the model affect the final prediction.

@@

@@

@@

%
%However, to understand how the model should
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{depTreeExample}
  \vspace{-6mm}
 \caption{
Dependency tree provides valuable information that can help fix the prediction error.
In (a), the model mistakenly aligns the word green, which lead to wrong prediction.  
After examining the dependency tree (highlighted by pink squares), we can see the two \textbf{green} are attached to different words.
In (b), by editing the attention and forcing the two \textbf{green}'s alignment to be zero, the prediction label is corrected to \emph{neutral}. 
 }
\label{fig:depTreeExample}
 \vspace{-4mm}
\end{figure*}

### 2018-03-30 paper/discussion.tex
update application 6.1

To conclude, this work introduces the \textbf{NLIVZ}, a perturbation-driven visual interrogation system that provides the expert with a streamlined exploration environment for testing their hypothesis and obtain intuition about the model. The proposed system frees the researchers from interruptions and tedious operations and allows them to focus on the more productive activities.

%\begin{itemize}
%    \item how the visualization approaches generalizes to other NLP task
%    \item limitation
%    \item future direction
%\end{itemize}

>>>>

To conclude, this work introduces a perturbation-driven visual interrogation system that provides the expert with a streamlined exploration environment for testing hypothesis and obtain intuition about the model. The proposed system frees the researchers from interruptions and tedious operations, therefore, allows them to focus on the more productive activities.

### 2018-03-30 paper/evaluation.tex
update application 6.1

\section{Evaluation and Feedback}
The motivation for designing the proposed tool is driven by the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we work closely with two NLP experts, and their constant evaluation and feedback have helped shape the tool we see today. 
%
Since these two NLP experts are heavily involved in the design process, they may not be the best candidates for identifying potential issues of the tool due to familiarity.  To uncover the limitation and identify areas for improvements, we gathered a wider audiences from either visualization (4 researchers include 3 Ph.D. level student and one postdoc researcher) or NLP background (5 Ph.D. level students who are familiar with the concept of natural language inference and attention) for obtaining feedback for the tool.

With the goal to identify the potential interface design issues, we first conduct an informal demonstration-and-feedback session with the visualization group. We start by explaining the basic natural language inference concept. We then demonstrate the features of the tool in detail while answering questions and address comments.
%
From this session, we have gathered complaints and suggestions on various aspect of the visual interface. Some of the identified issues include: (1) difficult to distinguish different type of predictions (distinct colors are added for the final version); (2) hard to recognize ground truth label (now, we use a green rectangle to indicate the ground truth); (3) lack of legends to understand the key elements in plots (legends are added). We address these interface issues before presenting the final version to the NLP group.

We conduct individual revaluation session with each participant from the NLP group, in which the participant is given 30 minutes to experiment with the tool after an overall feature demonstration. 
%
Since both the matrix and graph-based visual encoding is the most common attention representations used in NLP literature, most the participant can utilize them immediately and find the linked highlighting feature of the two views quiet useful.
% 
Once the participants familiarize with the tool, the first thing they try is often typing two similar sentences and examine the attention and prediction. After that, they will modified some words, or negate the hypothesis sentence, then check the attention and prediction again.
%
Such a ``perturb and observe'' operation demonstrates the most fundamental exploration strategy and matches well with the perturbation-driven paradigm the proposed tool aim to support.

@@

The participants also identify potential issues and places for improvement. 
One participant wishes the pipeline view could provide more detailed information compared to the current aggregated histogram visualization. 
Another participant suggests the possibility to examine multiple similar examples at a time (instead of one by one). 
%
Also, for the participants do not focus on NLI research, understand all the views at first can be a bit challenging. However, this issue can be addressed by the modular design, as the user can simply enable, only the sentence, attention, and prediction views, at first.
%
Overall, all the participants believe the proposed tool is very convenient for conduction experiment and exploring various hypothesis, which is essential to build intuitions about the model.

>>>>

\section{Evaluation and Feedback}
The motivation for designing the proposed tool is driven by the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we work closely with two NLP experts, and their constant evaluation and feedback have helped shape the tool we see today. 
%
Since these two NLP experts are heavily involved in the design process, they may not be the best candidates for identifying potential issues of the tool due to familiarity.  To uncover the limitation and identify areas for improvements, we gathered a wider audience from either visualization (4 researchers include 3 Ph.D. level students and one postdoc researcher) or NLP background (5 Ph.D. level students who are familiar with the concept of natural language inference and attention) for obtaining feedback for the tool.

With the goal to identify the potential interface design issues, we first conduct an informal demonstration-and-feedback session with the visualization group. We start by explaining the basic natural language inference concept. We then demonstrate the features of the tool in detail while answering questions and address comments.
%
From this session, we have gathered complaints and suggestions on various aspect of the visual interface. Some of the identified issues include: (1) difficult to distinguish different type of predictions (distinct colors are added to the final version); (2) hard to recognize ground truth label (now, we use a green rectangle to indicate the ground truth); (3) lack of legends to understand the key elements in plots (legends are added). We address these interface issues before presenting the final version to the NLP group.

We conduct individual revaluation session with each participant from the NLP group, in which the participant is given 30 minutes to experiment with the tool after an overall feature demonstration. 
%
Since both the matrix and graph-based visual encoding is the most common attention representations used in NLP literature, most the participant can utilize them immediately and find the linked highlighting feature of the two views quiet useful.
% 
Once the participants familiarize with the tool, the first thing they try is often typing two similar sentences and examine the attention and prediction. After that, they will modify some words, or negate the hypothesis sentence, then check the attention and prediction again.
%
Such a ``perturb and observe'' operation demonstrates the most fundamental exploration strategy and matches well with the perturbation-driven paradigm the proposed tool aim to support.

@@

The participants also identify potential issues and places for improvement. 
One participant wishes the pipeline view could provide more detailed information compared to the current aggregated histogram visualization. 
Another participant suggests the possibility to examine multiple similar examples at a time (instead of one by one). 
%
Also, for the participants do not focus on NLI research, understand all the views at first can be a bit challenging. However, this issue can be addressed by the modular design, as the user can simply enable, only the sentence, attention, and prediction views, at first.
%
Overall, all the participants believe the proposed tool is very convenient for conduction experiment and exploring various hypothesis, which is essential for building intuitions about the model.
